The Markov link method: a nonparametric approach to combine observations from multiple experiments

### 2018-08-22 paper/paper.tex
another abstraact revision

\begin{abstract}
We propose the Markov link method (\MLM), a way to synthesize experiments based on several measurement procedures.  For example, we might have different experiments measuring different aspects of human neurons, such as morphology, transcriptomics, or electrophysiology.  The \MLM{} uses this data find the  link between different aspects.  This link answers questions such as: ``We have observed the transcriptome of this cell -- what might its morphology have been?'' The method inspects how different \emph{measurement} procedures relate by using a collection of \emph{sample selection} procedures.   In doing so, it makes no assumptions about what the link looks like.  Instead, the \MLM{} relies on a conditional independence assumption.  We prove conditions under which the \MLM{} estimators are consistent, and simulations show that it gives accurate measures of uncertainty.  We evaluate the MLM on a pair of single-cell RNA techniques and gain new insight about the link between the two methods; we see aspects in which two different notions of `cell-type' are aligned, aspects in which they are not, and aspects about which we are unsure.  The \MLM{} suggests specific additional experiments which would help refine our understanding of the relationship between the two measurement procedures.
\end{abstract}

>>>>

\begin{abstract}
The Markov link method (\MLM) synthesizes experiments based on different measurement procedures.  For example, we might have different experiments measuring different aspects of human neurons, such as morphology, transcriptomics, or electrophysiology.  To combine knowledge across these experiments, we develop the idea of a ``link'' between measurement procedures.  This link should answer questions such as: ``We have observed the transcriptome of this cell -- what might its morphology have been?'' Without joint measurement of the transcriptome and morphology for the same neurons, it is challenging to estimate the link.  However, it turns out the problem is not impossible.  The \MLM{} estimates these links without joint measurement and without any assumptions about what the link looks like. Instead, the \MLM{} relies on a conditional independence assumption.  The estimation problem arising from this assumption is challenging, due to nonidentifiability problems; we give a rigorous characterization of the issue and show that the \MLM{} handles it correctly.  We prove conditions under which the \MLM{} estimators are consistent, and simulations show that it gives accurate measures of uncertainty.  We evaluate the \MLM{} on a pair of single-cell RNA techniques and gain new insight about the link between the two methods; we find that the two different notions of `cell-type' are generally well-aligned, with a few exceptions.  The \MLM{} also shows that some aspects of the link simply cannot be determined from the data we have.  The \MLM{} suggests specific additional experiments which would help refine our understanding of the relationship between the two measurement procedures.
\end{abstract}
